Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-24 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
In another word, the set of all quantum probability measures is one-to-one
corresponding to the set of states so that studying the set of states
is sufficient to understand all possible phenomenon regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties extending Gleason's theorem.
In particular, there is a QIVPM such that no state can be consistent
with it on every projector. However, if we only focus on the Born
QIVPMs~$\bar{\mu}$, that is, QIVPM's induced by a state~$\rho$
according to the Born rule, we find that if we monotonically shrink
the length of intervals, the states consistent with $\bar{\mu}$ will
be closer and closer to $\rho$, and vise visa. In another word, although
we can prove a finite precision version of Gleason's theorem, there
might be more rich phenomenon regarding to quantum probability when
considering finite precision measurement. Our result also highlight
the necessity to develop a mathematical theory of quantum mechanics
that intrinsically consistent with the resources available for realistic
measurement.

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As in
Table~\ref{tab:I_n}, To select the correct interval from these overlapping
intervals~$[\ell,r]$, we define the non-overlapping interior part
of a interval, $\textrm{Int}\left([\ell,r]\right)$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$ and their interior.
Although intervals~$[\ell,r]$ are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping. See Table~\ref{tab:I_2} for a concrete example
when $n=2$.}

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.(\ref{eq:QIVPM-constraints}), the condition really need to be
checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto the subspace which the other one projects
onto.
\end{itemize}
\noindent No matter which case, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

As a concrete example of Thm.~\ref{thm:Finite-precision-Born}, when
$n=2$, two states~$\rho_{0}=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{1}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$
can induce two QIVPMs~$\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$. Given any projector
$P$ in the minimal subspace of events containing $\proj{0}$, $\proj{1}$,
and $\proj{2}$, $\events_{C}$, the interval-valued probability of
the induced QIVPMs, $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right)$,
is computed straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ based on $\mu_{\rho_{0}}^{\mathrm{B}}$
and $\mu_{\rho_{1}}^{\mathrm{B}}$ on $\events_{C}$.}

@@

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the states consistent with it on every project, $\coreBorn\left(\bar{\mu},\events\right)$,
has a unique element. \end{thm}

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we introduce a few technical notions. We call a state~$\rho$
a \emph{totally mixed state} if $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.
We also define a notion of distance between two states~$\rho$ and
$\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199},
and the set of all states whose distance with $\rho$ is bounded by
$\epsilon$ is denoted by $B_{\epsilon}\left(\rho\right)=\set{\rho'}{\left\Vert \rho-\rho'\right\Vert \le\epsilon}$.

In another word, if a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $1/n$. Reversely, if the distance between $\rho$ and
$\rho'$ is bounded by $\frac{1}{12n}$, then $\rho'$ is consistent
with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.

@@

\noindent For example, $\rho_{0}$ and $\rho_{1}$ are both totally
mixed states. Since $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{24}$,
we should have $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the state very close to
$\rho$ are guaranteed to belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.

>>>>

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
In another word, the set of all quantum probability measures is one-to-one
corresponding to the set of states so that studying the set of states
is sufficient to understand all possible phenomenon regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties extending Gleason's theorem.
If we only focus on the Born QIVPMs~$\bar{\mu}$, that is, QIVPM's
induced by a state~$\rho$ according to the Born rule, we find that
if we monotonically shrink the length of intervals, the states consistent
with $\bar{\mu}$ will be closer and closer to $\rho$, and vise visa.
However, there is a QIVPM such that no state can be consistent with
it on every projector. In another word, although the Born rule and
Gleason's theorem can be extended to consider finite precision measurement,
the problem of finite precision measurement itself is not limited
in just applying error analysis in experimental results, but it highlight
some new phenomenon regarding to the foundation of quantum machines,
and it is necessary to embedding the idea of the resources available
for realistic measurement into the formulation of quantum theory.

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As in
Table~\ref{tab:I_n}, to select the correct interval from these overlapping
intervals~$[\ell,r]$, we define the non-overlapping interior part
of a interval, $\textrm{Int}\left([\ell,r]\right)$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$ and their interior.
Although intervals~$[\ell,r]$ are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping. See Table~\ref{tab:I_2} for a concrete example
when $n=2$.}

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.~(\ref{eq:QIVPM-constraints}), the condition really need to
be checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace which the other one projects
onto.
\end{itemize}
\noindent No matter which case, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

As a concrete example of Thm.~\ref{thm:Finite-precision-Born}, when
$n=2$, two states~$\rho_{0}=\frac{1}{3}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{1}{3}\cdot\proj{2}$
and $\rho_{1}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$
can induce two QIVPMs~$\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$. Given any projector
$P$ in the minimal subspace of events containing $\proj{0}$, $\proj{1}$,
and $\proj{2}$, $\events_{C}$, the interval-valued probability of
the induced QIVPMs, $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right)$,
can be computed straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ based on $\mu_{\rho_{0}}^{\mathrm{B}}$
and $\mu_{\rho_{1}}^{\mathrm{B}}$ on $\events_{C}$.}

@@

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the states consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
has a unique element. \end{thm}

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we introduce a few technical notions. We call a state~$\rho$
a \emph{totally mixed state} if $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.
For example, $\rho_{0}$ and $\rho_{1}$ are both totally mixed states,
but any pure state is not totally mixed. We also define a notion of
distance between two states~$\rho$ and $\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199},
and the set of all states whose distance with $\rho$ is bounded by
$\epsilon$ is denoted by $B_{\epsilon}\left(\rho\right)=\set{\rho'}{\left\Vert \rho-\rho'\right\Vert \le\epsilon}$.
For example, since $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{24}$,
we have $\rho_{0}\in B_{\epsilon}\left(\rho_{1}\right)$ and $\rho_{1}\in B_{\epsilon}\left(\rho_{0}\right)$.

After we have enough definitions, we can replace the QIVPM~$\bar{\mu}$
in Thm.~\ref{cor:Gleason's} by a Born QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
and estimate the states consistent with it by the following theorem.

@@

\noindent In another word, if a state~$\rho'$ is consistent with
$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$, its distance with $\rho$,
$\left\Vert \rho-\rho'\right\Vert $, must be bounded by $1/n$. Reversely,
if the distance between $\rho$ and $\rho'$ is bounded by $\frac{1}{12n}$,
then $\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.
For example, we should have $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the state very close to
$\rho$ are guaranteed to belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.
